---
title: "Using Vector Embeddings For Sentiment Analysis"
author: "Rod Acosta, Kevin Furbish, Ibrahim Khan, Anthony Washington"
format:
  html:
    code-fold: true
course: STA 6257 - Advanced Statistical Modeling
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

## 1. Introduction

Sentiment Analysis (SA) is a branch of Natural Language Processing (NLP) that computationally extracts the “treatment of opinions, sentiments and subjectivity” [@MEDHAT20141093] of digital text. 

### 1.1 Sentiment Analysis
Sentiment analysis, a crucial facet of natural language processing (NLP), involves interpreting and classifying emotions within text data. By understanding public opinion, businesses can refine marketing strategies, improve products, and enhance customer satisfaction. The rapid growth of social media and e-commerce platforms has amplified the importance of sentiment analysis, enabling real-time feedback and insights into consumer behavior.

The study by Hasan, Maliha, and Arifuzzaman (2019) demonstrates the potential of Twitter data for sentiment analysis, highlighting the application of NLP frameworks to gauge public opinion on products. Their methodology integrates the Bag of Words (BoW) and Term Frequency-Inverse Document Frequency (TF-IDF) models, coupled with logistic regression, to classify tweets as positive or negative with an accuracy of 85.25%. This approach underscores the benefits of combining BoW and TF-IDF to enhance sentiment analysis precision.

In a similar vein, Kathuria, Sethi, and Negi (2022) explored sentiment analysis on e-commerce reviews, employing machine learning (ML) models such as logistic regression, AdaBoost, SVM, Naive Bayes, and random forest. Utilizing the Women’s E-commerce Clothing Reviews dataset, they analyzed review texts, ratings, and recommendations to understand consumer behavior. Their findings underscore the significance of electronic word-of-mouth (eWOM) in shaping customer attitudes and product sales, providing e-commerce businesses with actionable insights to improve marketing strategies and customer satisfaction.

The foundational work by Nasukawa and Yi (2003) emphasizes a more granular approach to sentiment analysis, focusing on extracting sentiments linked to specific subjects rather than entire documents. Their prototype system, employing a syntactic parser and a sentiment lexicon, achieves high precision in detecting sentiments in web pages and news articles. By concentrating on local sentiment expressions, their method offers detailed insights into specific opinions, aiding businesses in monitoring public opinion and addressing unfavorable sentiments effectively.

Collectively, these studies highlight the evolution and application of sentiment analysis using NLP, illustrating its critical role in extracting valuable insights from vast amounts of text data. By leveraging advanced NLP techniques and ML models, businesses can gain a deeper understanding of consumer sentiment, thereby enhancing their strategic decision-making processes.

### 1.2 Vector Embeddings In Natural Language Processing
One technique for encoding or describing the sentiment of a word or group of words is to use vector embeddings (or embeddings). Before discussing the use of embeddings for SA, it’s important to first understand what embeddings are. Embeddings are said to be “one of the most important topics of interest” in NLP in the last decade [@camacho2020embeddings].

An early way to represent a word for NLP systems was to have a vocabulary of words, and each word is represented as an atomic, simple index into the vocabulary [@mikolov2013word2vec]. Embeddings are an alternative method to encode natural language as a vector representation. Embeddings can be used to represent words [@bengio2000neural]; [@collobert2011natural]; [@pennington2014glove], sentences [@kiros2015skip] and documents [@pilehvar2020embeddings], as well as to encode more specific meanings of natural language for tasks such as sentiment analysis, topic categorization and word sense disambiguation [@pilehvar2020embeddings]. 

The use of word embeddings in NLP is an improvement over representing words as an index into a vocabulary since embeddings can encode relationships or similarity of words. For example, when using a simple index into a vocabulary to represent a word, “boy” would have one index and “man” would have another index, but there is no indication that these two words are related. Vocabulary indexes also fail to represent that a word may have multiple meanings. For example, “mouse” would be one entry in a vocabulary, but would fail to indicate if the word refers to an animal, or computer input device [@camacho2020embeddings].

Word embedding models can capture fairly detailed semantic and syntactic patterns, however how these patterns are encoded in a vector is often unclear [@pennington2014glove]. Unlike indexes into a vocabulary, word embeddings have the advantage of being able to be compared for similarity. Comparison is often done either by measuring distance between vectors, or the angle between vectors [@pennington2014glove]. [@mikolov2013word2vec] discovered in their famous “Word2Vec” model that simple vector arithmetic over their embedding model allowed evaluation of analogies. For example, $vector(king) - vector(man) + vector (woman)$ resulted in a vector that was closest to the vector for “queen”.

Embedding models also exist for more complicated NLP problems than word representations. Sentence embedding models can be built on top of word embeddings models. This is an improvement over using “bag of words” representations for sections of natural language longer than a single word. Bag of word models often use vectors of one-hot encoding, however these representations have very high dimensionality and sparseness. These issues led researchers to look for alternatives and they found embeddings to be a valuable technique [@pilehvar2020embeddings]. Sentence embeddings compose word embeddings for each of the words of a sentence into a single vector that encodes semantic and syntactic properties of the sentence and can be compared for similarity across those properties [@kiros2015skip]. As with using bag of words to encode sentences, bag of word models are similarly poor representations of documents for many NLP tasks. In such cases, encoding the document into a document embedding is a useful alternative representation [@pilehvar2020embeddings]. Since bag of word models ignore word ordering, they are a problematic representation for many NLP tasks, including sentiment analysis [@pilehvar2020embeddings], which will be the focus for the rest of this work.

### 1.3 Using Embeddings For Sentiment Analysis
Now that we have touched on the topics of sentiment analysis and embeddings, let's dive into the use of embeddings for sentiment analysis. The performance of natural language processing (NLP) tasks including question answering & sentiment analysis, has significantly improved due to word embedding and deep learning models over the years [@10255396]. Ultimately our goal when converting text into embeddings is to be able to extract the meaning of the words in the text and allow our model to learn from it. There are two very popular word embedding methods, of which we have already touched upon slightly, the two methods are Word2Vec and Global Vectors (GloVe). Using one of these two methods along with an embedding model we will be able to categorize data into positive, negative, or neutral sentiments by evaluating the overall polarity of the input data. This branch of NLP has been increasingly becoming more and more popular, specially for processing and gathering the sentiment of posts on social media sites, forums, an the web in general. 

Our two popular pretrained word  methods were each trained on different datasets. Word2Vec was trained on a portion of the Google News dataset with 300-dimensional vectors and GloVe was trained on 6 billion tokens, 400k vocabulary words, and 100-dimensional vectors from Wikipedia 2014 [@10543279]. The drawback to most embedding based models is that they typically only utilize pre-trained word embeddings that can not capture the effective information in the text or input,  such as the contextual sentiment information of both targets and aspects [@9397340]. This is problematic because this leads words with similar contexts but opposite sentiment polarity (good and bad for example) to be mapped to neighboring word vectors making our models less accurate[@7296633]. We will be tackling this issue more in depth in this paper and exploring other forms of embeddings for sentiment analysis and models along the way!


## 2. Methods
```{r}
install.packages('reticulate', repos = "http://cran.us.r-project.org")
install.packages('tensorflow', repos = "http://cran.us.r-project.org")
install.packages('keras', repos = "http://cran.us.r-project.org")
install.packages('dplyr', repos = "http://cran.us.r-project.org")
install.packages('text2vec', repos = "http://cran.us.r-project.org")
install.packages('tidyr', repos = "http://cran.us.r-project.org")
```

```{r}
reticulate::install_python()
library(tensorflow)
install_tensorflow(envname="r-tensorflow")
tensorflow::as_tensor(1)
library(keras)
install_keras()
#install.packages('fs')

tensorflow::as_tensor(1)
library(text2vec)
library(dplyr)
library(tidyr)
```


```{r}
orig = read.csv("training.1600000.processed.noemoticon.csv", header=TRUE)
names(orig) <- c("polarity", "id", "date", "query", "user", "text")
```

```{r}
ds = rbind(head(orig, n=5000), tail(orig, n=5000)) # use small dataset in dev
# ds <- orig
max_features <- 5000
maxlen <- 500
vector_dimensions <- 7

# First create a vocabulary of all the individual words used in the dataset
it = itoken(ds$text, preprocessor = tolower, tokenizer=word_tokenizer)
vocab <- create_vocabulary(it) 
vocab_sorted <- vocab[order(vocab$term_count, decreasing=TRUE),] %>% mutate(idx = row_number())
vocab_sorted <- head(vocab_sorted, n=max_features-1) #take only the top most common words

# Now create a mapping from the vocabulary word to a integer
# representation for each word. Use factor indexing so the word
# can be looked up quickly and retrieve the mapped integer value.
vocab_translate = vocab_sorted$idx
names(vocab_translate) = as.factor(vocab_sorted$term)

```

```{r}

# Vectorize every tweet in the dataset by looking up the integer index
# from the vocabulary associated with each word in the tweet.

vectorize_row <- function(r) {
  tokenized = word_tokenizer(r)[[1]]
  vectorized <- c(sapply(tokenized, function(x) vocab_translate[x]), rep(0,maxlen-length(tokenized)))
  return(vectorized)
}

tokenized_ds <- sapply(ds$text, vectorize_row)
tokenized_ds <- t(tokenized_ds) # transpose since sapply returns the tweets in columns instead of rows
tokenized_ds <- replace_na(tokenized_ds, 0) # any word outside our top vocab is NA, so replace with 0 for unknown word
dimnames(tokenized_ds) <- list(NULL, NULL) # remove the tweets as indexes into the matrix

set.seed(42)
sample_size <- floor(0.7*nrow(tokenized_ds)) # 70% training set
train_idx <- sample(seq_len(nrow(tokenized_ds)), size=sample_size)
train_x <- tokenized_ds[train_idx, , drop=FALSE]
train_y <- ds$polarity[train_idx]
test_x <- tokenized_ds[-train_idx, , drop=FALSE]
test_y <- ds$polarity[-train_idx]
```

```{r}
embedding_model <- keras_model_sequential() %>%
  layer_embedding(input_dim = max_features, output_dim = vector_dimensions,  
                  input_length = maxlen) %>%
  layer_flatten() %>%
  layer_dense(units = 1, activation = "sigmoid")
embedding_model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("acc")
)
summary(embedding_model)
```

```{r}
emb.history <- embedding_model %>% keras::fit(
  train_x, train_y, 
  epochs = 6,
  batch_size = 32,
  validation_split = 0.2
)
```

```{r}
emb.results <- embedding_model %>% evaluate(test_x, test_y)
```







## 3. Analysis and Results
### 3.1 Dataset Description
For sentiment analysis, we will be using a data set that contains computer generated tweets for training data and another collection of tweets for testing. Any tweet with positive emoticons was assumed to be positive, and any tweet with negative emotions was assumed to be negative. The emoticons were removed from the dataset. The dataset was last updated 4 years ago and was posted by Abhishek Shrivastava. The computer-generated tweets that make up the training data are organized into ten columns. The first column is textID used to uniquely identify each tweet in the dataset. The second column is the text which is the entire tweet used. The third column is the selected text of the tweet used to perform sentiment analysis. The fourth column is the labeled sentiment of the tweet, which is classified as neutral, negative, or positive. The fifth column is the time of day the tweet was posted which is classified as morning, noon, or night. The sixth column stores the age range that the user is in. The seventh column is the country the tweet was posted from when the user posted. The eighth column is the given population of the country. The ninth column is the given land area of the country. The tenth column is the given population density of the country. During the literature review process, twitter datasets were often used for sentiment analysis such as “Sentiment Analysis with NLP on Twitter Data” by (Hasan et al. 2019) and “Natural Language Processing for Sentiment Analysis An Exploratory Analysis on Tweets” by (Chong et al. 2014). Therefore, this dataset of generated tweets seems appropriate for our sentiment analysis research project.

### Data and Visualization


### Statistical Modeling

### Conclusion

## References
Khurana, D., Koli, A., Khatter, K. et al. Natural language processing: state of the art, current trends and challenges. Multimed Tools Appl 82, 3713–3744 (2023). https://doi.org/10.1007/s11042-022-13428-4

Jan Sawicki, Maria Ganzha, Marcin Paprzycki; The State of the Art of Natural Language Processing—A Systematic Automated Review of NLP Literature Using NLP Techniques. Data Intelligence 2023; 5 (3): 707–749. doi: https://doi.org/10.1162/dint_a_00213

M. R. Hasan, M. Maliha and M. Arifuzzaman, "Sentiment Analysis with NLP on Twitter Data," 2019 International Conference on Computer, Communication, Chemical, Materials and Electronic Engineering (IC4ME2), Rajshahi, Bangladesh, 2019, pp. 1-4, doi: 10.1109/IC4ME247184.2019.9036670. keywords: {Sentiment analysis;Twitter;Load modeling;Data mining;Analytical models;Companies;Natural language processing (NLP);Twitter;data mining;Sentiment analysis},

W. Y. Chong, B. Selvaretnam and L. -K. Soon, "Natural Language Processing for Sentiment Analysis: An Exploratory Analysis on Tweets," 2014 4th International Conference on Artificial Intelligence with Applications in Engineering and Technology, Kota Kinabalu, Malaysia, 2014, pp. 212-217, doi: 10.1109/ICAIET.2014.43. keywords: {Sentiment analysis;Twitter;Semantics;Feature extraction;Machine learning algorithms;Tagging;Sentiment Analysis; Natural Language Processing; Tweets},

