---
title: "Analysis on Life Expectancy using Bayesian Linear Regression"
author: "Kayla Mota, Mary Catherine Morrow, Summer Dahlen, Shijie Geng"
date: '`r Sys.Date()`' 
format:
  html:
    code-fold: true
course: STA 6257 - Advanced Statistical Modeling
bibliography: references.bib # file contains bibtex for references
#always_allow_html: true # this allows to get PDF with HTML features
self-contained: true
execute: 
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---


Packages & Library
```{r}
# Install and load required packages
#install.packages("readxl")  # Install if not already installed
#install.packages("corrplot")  # Install if not already installed
library(readxl)
library(corrplot)

# install.packages("rstanarm") 
library(rstan)
library(rstanarm)
library(ggplot2)
library(bayesplot)

# this option uses multiple cores if they're available
options(mc.cores = parallel::detectCores())
```


## 1. Introduction

Bayesian Linear Regression (BLR) is a conditional data analytical method
for understanding the relationship between independent and dependent
variables based on Bayes Theorem while handling uncertainties with
accuracy and interpretability[@walker2007application]. In the prevalent
perspective, the primary method for estimating parameters in a linear
regression model is the frequentist approach. However, this methodology
defines the probability of uncertain events as the relative frequency of
their occurrence[@hajek1996mises], and the resulting model parameters
are derived exclusively from the observed data [@hajek2009fifteen].
Additionally, the frequentist approach may distort parameter estimates
and affect the validity of statistical tests when dealing with a small
datatset [@zyphur2015bayesian]. Unlike frequentist method, BLR is a more
flexible statistical model in that it captures all uncertain variables
random [@klauenberg2015tutorial] and is also known to be used for
experiments that have small sample sizes [@sugiarto2021determining] and
various data types [@kruschke2010bayesian]. Moreover, BLR eliminates the
need for p-values, which offers richer information on parameters, and
allows for simultaneous single change point detection in a multivariate
sample [@seidou2007bayesian].

The BLR is characterized by viewing parameters as random variables,
introducing the concept of prior, likelihood, and posterior
distributions [@permai2018linear; @chen2024positional]. The BLR method
involves multiple steps: predicting prior values, imputting collected
data, and obtaining probability in the form of a distribution. This
distribution, called the posterior, is proportional to the prior
distribution multiplied by the likelihood function
[@baldwin2017introduction]. In BLR, even when utilizing representative
or informed priors, it is crucial to perform a sensitivity analysis to
assess the impact of different prior specifications on the posterior
results. This aims to ensure that the findings are not dependent on the
choice of prior [@kruschke2021bayesian]. Once the experiment is
conducted and the posterior distribution is obtained, the next step is
to conduct a check on skewness of the distributions, as well as
heteroscedasticity. Although scale mixtures of normal are often used in
regression analysis to capture heteroscedasticity and outliers,
skew-symmetric scale mixtures of normal (SSSM) can capture asymmetry,
making it more robust to use in BLR [@rubio2016bayesian]. The posterior
distribution is where the Markov Chain Monte Carlo (MCMC) method
involves simulating random draws from the posterior distribution,
allowing for exploration of the parameter space and estimation of the
posterior distribution even in complex, high-dimensional models
[@robert2018accelerating; @rojas2020lose]. Confirming the convergence of
MCMC chains is important for the reliability of the results. Evidence of
convergence provides confidence in the validity of the
samples[@marcotte2018gibbs]. After obtaining the posterior results,
inferences can be used for the prior distribution of the current/next
experiment [@rubio2016bayesian].


To explain, posterior probabilities are assigned to the spectrum of
values that parameters can assume. The more prominent the peak, the more
effective it is as an estimate because a single value has a higher
probability compared to other values. In such cases, the credibility
interval, indicating a 95% range of probable parameter estimates, is
narrower. Conversely, when the peak is less pronounced, other estimates
are also plausible, resulting in a wider credibility interval for the
95% range of probable parameter estimates [@shetty2013evidence;
@gill2002bayesian].

BLR has been used in many disciplines due to its numerous strengths.
Researchers have the option to incorporate past findings as informative
priors, a practice observed in fields such as construction and
biomedicine. This analytical approach produces outcomes that amalgamate
previous results with the data from a present study, treating it as if
all existing datasets were collectively analyzed. An example of using
BLR includes determining correction coefficients to previously
inaccurate concrete mixture formulas to ensure the correct use and
amount of chemicals to prevent and then predict the breakdown of the
concrete [@zgheib2019bayesian]. Another example of BLR is re-determining
traffic-flow rate values for vehicles to be able to ensure optimal
traffic flow and minimize to a safe number of vehicles at intersections,
which allows for the previous rate values from over 20 years ago to be
updated [@sugiarto2021determining]. BLR can also be used to determine
the load and strain a bridge can take before it is unsafe to be used
[@zhang2022long], and it can be helpful to predict genetic values for
genomic selection for plants and animals to prevent the risk of passing
on genes that could cause illnesses [@perez2010genomic].

## 2. Methods

### 2.1 Linear Regression

Regression analysis is a statistical method that allows to examine the
relationship between two or more variables of interest. Suppose the
response variable $Y$ is a function of several predictor variables,
$x_{i1}, x_{i2},..., x_{ik}$.

To fit a model:

$$
Y = X \beta + \epsilon     \quad(1)
$$

where,

$Y = (y_1, y_2,..., y_n)$ is the n×1 vector of response; $X$ is a n×p
design matrix, $p = k+1$; $\beta = (\beta_0, \beta_1,...., \beta_{k})$
is the (k+1)×1 vector of parameters and
$\epsilon = (\epsilon_1, \epsilon_2,...., \epsilon_n)$ the n×1 vector of
errors, $\epsilon$ \~ $N_n(0, \sigma^2I)$.

The likelihood function of $y$ given $\beta$ and $\sigma^2$ is defined
as follows

$$
f(Y_i = y_i|x, \beta, \sigma^2) = f_y(y|x,\beta,\sigma^2) = (2\pi\sigma^2)^{-\frac{n}{2}} exp^{-\frac{1}{2\sigma^2}(y-x\beta)^T(y-x\beta)}  \quad(2)
$$

To make inference, the $\beta$ parameters are determined by maximizing
the likelihood function.

### 2.2 Bayesian Linear Regression

From the Bayesian perspective, linear regression is approached through
probability distributions. Instead of estimating the response variable,
y, as a single value, it's viewed as being drawn from a probability
distribution. In Bayesian Linear Regression, the model assumes that the
response follows a normal distribution, expressed as

$$
y \sim N(\beta^TX, \sigma^2I)  \quad(3)
$$

where,

y, which is assumed as a normal distribution centered by its mean and
variance. In linear regression, the mean is calculated by the transpose
of the weight matrix and the predictor matrix. The variance is
determined by squaring the standard deviation $\sigma$ and then
multiplying it by the identity matrix.

Additionally, the parameters in the model are generated from a
probability distribution. The posterior probability of these parameters
is proportional to the prior and likelihood [@minka2000bayesian].

$$
P(\beta|y, X) = \frac{P(y|\beta, X) × P(\beta|X)}{P(y|X)}  \quad(4)
$$

In Bayesian statistics, when the prior distribution and the posterior
distribution belong to the same family of probability distributions,
they are called conjugate distributions. Using the Bayes' theorem, we
have

$$
f(\beta, \sigma^2|y,x) = \frac{f_y(y|x, \beta,\sigma^2)f(\beta|\sigma^2,m,V)f(\sigma^2|a,b)}{f_y(y)}  \quad(5)
$$

The equation can be broken down as,

-   $f(\beta, \sigma^2|y,x)$: This is the joint posterior distribution
    of the coefficients $\beta$ and the error variance $\sigma^2$ given
    the data $y$ and the predictors $x$

-   $f_y(y|x, \beta,\sigma^2)$: This is the likelihood function which
    describes the probability of observing the data $y$ given the
    predictors $x$ and the parameters $\beta$ and $\sigma^2$.

-   $f(\beta|\sigma^2,m,V)$: It represents the prior distribution for
    coefficients $\beta$ given the error variance $\sigma^2$, with mean
    vector $m$ and covariance matrix $V$.

-   $f(\sigma^2|a,b)$: It represents the prior distribution for the
    error variance $\sigma^2$, which is assumed to be an inverse gamma
    distribution with parameter $a$ and $b$.

-   $f_y(y)$: It is the marginal likelihood and often treated as a
    constant.

A non-conjugate prior refers to a prior distribution that does not
belong to the same family of distributions as the posterior distribution
after observing data. In such cases, the posterior distribution can be
expressed as

$$
f(\beta|y, X, \sigma^2) ∝ f_y(y|X, \beta)\prod_{i=1}^{n}f(\beta_i|m_i,v_i)  \quad(6)
$$

-   $f(\beta|y, X, \sigma^2)$: This is the posterior distribution of
    coefficients $\beta$ given the dependent variable $y$, independent
    variable $X$ and error variance $\sigma^2$.

-   $f_y(y|X, \beta)$: This is likelihood function, denoting the
    probability of the observing data $y$ given independent variable $X$
    and the coefficient parameters $\beta$.

-   $f(\beta_i|m_i,v_i)$: This is prior distribution of each coefficient
    $\beta_i$ with mean $m_i$ and variance $v_i$.

## 3. Analysis and Results

### 3.1 Data Introduction

The selected life expectancy dataset from the World Health Organization
(WHO) comprises of 21 health, economic, and social factors taken from
179 countries between years 2000 and 2015. Below is a description of the
variables:

-   **Country:** Categorical, one of 179 countries included in the
    dataset specific to data collection location.

-   **Region:** Categorical, one of 9 regions specific to data
    collection location.

-   **Year:** Continuous, range from 2000 to 2015.

-   **Infant Deaths (InfDth):** Continuous, per 1000 population.

-   **Under_five_deaths (UnFiveDths):** Continuous, number of deaths of
    children under 5 years of age per 1000 population.

-   **Adult Mortality (AdultMort):** Continuous, rate of deaths of
    adults between 15 and 60 years old, per 1000 population.

-   **Alcohol_consumption:** Continuous, recorded per capita (15+ year
    olds) consumption in litres of pure alcohol.

-   **Hepatitis B (HepB):** Continuous, percentage of Hepatitis B (HepB)
    immunization coverage among 1-year-old.

-   **Measles:** Continuous, percentage of coverage of Measles
    containing vaccine first dose (MCV1) immunization among 1-year-old.

-   **BMI:** Continuous, average body mass index (BMI) of entire
    population; measure of nutritional status in adults.

-   **Polio :** Continuous, percentage of Polio (Pol3) immunization
    coverage among 1-year-old.

-   **Diphtheria :** Continuous, percentage of diphtheria tetanus toxoid
    and pertussis (DTP3) immunization coverage among 1-year-old.

-   **Incidents_HIV (HIV):** continuous, incidents of HIV per 1000
    population, ages 15-49.

-   **GDP_per_capita (GDP):** Continuous, GDP per capita in USD.

-   **Population_mln (Population):** Total population in millions.

-   **Thinness_ten_nineteen_years (ThinAd):** Continuous, prevalence of
    thinness among children and adolescents for ages 10 to 19, %.

-   **Thinness_five_nine_years (ThinChild):** Continuous, prevalence of
    thinness among children for ages 5 to 9, %.

-   **Economy_status_Developed:** Categorical, Binomial; Developed
    Country.

-   **Economy_status_Developing:** Categorical, Binomial; Developing
    Country.

-   **Schooling:** Continuous, average years that people aged 25+ spent
    in formal education.

-   **Life_expectancy (LifeExp):** Continuous, average of the population
    of both genders in years.

#### 3.1.1 Load the dataset and inspect the first five records.

```{r, warning=FALSE, echo=T, message=FALSE}
# loading the packages required for data cleaning and BLR analysis.
library(tidyverse)
library(gtsummary)
library(skimr)
library(randomForest)
library(GGally)
library(corrplot)
library(car)
library(ggrepel)
library(rstan)
library(rstanarm)
library(bayesplot)


library(readxl)
ledata <- read_excel("C:/Users/kayla/Downloads/Life-Expectancy-Data-Updated.xlsx", col_names = FALSE)

names(ledata) <- c("Country", 'Region',"Year", "InfDth","UnFiveDths","AdultMort", "Alcohol","HepB","Measles","BMI","Polio", "Diphtheria", "HIV", "GDP", "population","ThinAd","ThinChild","Schooling",'Developing','Developed', "LifeExp")
str(ledata)

```

### 3.2 Exploration Data Analysis

#### 3.2.1 Summarize the Dataset

Table 1 below describes each continuous variable from the life
expectancy dataset by region.

```{r, warning=FALSE, echo=T, message=FALSE}
ledata %>% select(-c('Country', 'Year')) %>%
  tbl_summary(
    by = Region,
    label = list(InfDth ~ "Infant Deaths, per 1000",
                 UnFiveDths ~ "Deaths under 5 years of age",
                 AdultMort ~ "Adult Mortality Probability",
                 Alcohol ~ "Alcohol Consumption (L)",
                 HepB ~ "Hepatitis B Immunization (%)",
                 Measles ~ "Measles Immunization (%)",
                 BMI ~ "Average Body Mass Index", 
                 Polio ~ "Polio immunization (%)", 
                 Diphtheria ~ "Diphtheria Tetanus Toxoid and Pertussis immunization (%)", 
                 HIV ~ "Incidents of HIV per 1000",
                 GDP ~ "Gross Domestic Product (USD)",
                 population ~ "Population, in millions", 
                 ThinAd ~ "Thinness among 10-19 years (%)",
                 ThinChild ~ "Thinness among 5-9 years (%)",
                 Schooling ~ "Average Years of Schooling",
                 Developed ~ "Developed Country",
	               Developing ~ "Developing Country",
                 LifeExp ~ "Life Expectancy (yrs)"),
    missing_text = "(Missing)",
    statistic = list(
      all_continuous()~ "{mean} ({sd})",
      all_categorical()~ "{n}/{N} ({p}%)"
    )
  ) %>%
  add_n()%>%
  add_overall()%>%
  modify_header(label="**Variable**") %>%
  modify_caption("**Table 1. Dataset Summary**") %>%
  modify_footnote(all_stat_cols()~ "Mean (SD)") %>%
  bold_labels()
```

#### 3.2.2 Verify the correlation among variables.

```{r}
# Checking the correlation between those continuous variables and life expectancy.
# Remove these categorical variables from dataset.
ledata_continuous <- ledata %>% select(-c('Country', 'Region', 'Year', 'Developing', 'Developed'))

# Remove NA values from ledata_continuous
ledata_noNA <- na.omit(ledata_continuous)
#co_ledata <- ledata_noNA %>% select(-c('Country', 'Region', 'Year'))

# Compute correlation of continuous variables and life expectancy
cor(ledata_noNA)

# Visualize correlation matrix

ggcorrplot::ggcorrplot(round(cor(ledata_noNA),2), type = 'lower', title = 'Correlation within the life Expectancy dataset', show.legend = TRUE, lab = TRUE, lab_size = 2)
```

After conducting a correlation analysis, we identified that the
variables adult mortality, deaths under five years old and infant deaths
each have a strong negative correlation with life expectancy. The
variable population is not related to life expectancy at all. We
consider setting threshold of correlation between \[-1.00 \< -0.55\] and
\[0.55 \< 1.00\]. Any variables' correlation with response between
\[-1.00 \< -0.55\] and \[0.55 \< 1.00\] can be included for Bayesian
Linear Regression. Thus, GDP, BMI, Dephtheria, Polio, schooling, and HIV
also can be used to predict life expectancy.

#### 3.2.3 Multicollinearity Examination

```{r}
# Build a ordinary linear model for subsequent vif check
olr <- lm(LifeExp ~., data = ledata_noNA)

# Check Multicollinearity using vif()
vif(olr)

# Extract selected variables for subsequent BLR analysis
BLR_data <- ledata_noNA %>% select(c(GDP, BMI, Alcohol, Schooling, HIV,LifeExp))
```

Through checking the multicollinearity among these variables, the vif
values of infant deaths, deaths under five years old, adult mortality,
Polio, Diphtheria, thinness among those between 10 and 19, and thinness
among children for ages 5 to 9 are greater than 5, indicating
significant multicollinearity between them. Thus, given the results of
correlation and multicollinearity checking, we select the independent
variables of GDP, BMI, alcohol, schooling and HIV and dependent variable
of life expentency to contribute to our final dataset used for BLR
analysis.

#### 3.2.4 Data Visualization

Life expectancy follows a distribution skewed to the left, with the peak
occurring at 70-75 years for the highest life expectancy.

```{r, warning=FALSE}
hist(ledata_noNA$LifeExp, 
     main = "Figure 3. Distribution of Life Expectancy (years)",
     xlab = "Life Expectancy",
     border = "white", 
     col = "lightblue",
     ylim = c(0, 800))
```

The line graph illustrates the life expectancy trends across various
regions over the years. There has been a consistent rise in life
expectancy across all regions since 2000. North America and European
Union exhibit the highest life expectancy, while Africa shows the lowest
among these regions.

```{r, warning=FALSE, message=FALSE}
# Calculate max, mean, and min life expectancy by year and region
life_expectancy_summary <- ledata %>%
  group_by(Year, Region) %>%
  summarise(
    mean_life_expectancy = mean(LifeExp) )

# Create a line plot for max, mean, and min life expectancy in all regions over the years
ggplot(life_expectancy_summary, aes(x = Year)) +
  geom_line(aes(y = mean_life_expectancy, color = Region), linetype = "solid") +
  labs(title = "Figure 4. Regional Life Expectancy by Year (years)", x = "Year", y = "Life Expectancy", color = "Region") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") 

```

The graphs depict the correlation between schooling and life expectancy
across different regions. Generally, life expectancy tends to increase
with higher levels of schooling. Comparatively, more developed regions
such as North America, the European Union, and the Rest of Europe
exhibit both higher average schooling years and life expectancy.

```{r}
ledata_noNA %>% ggplot(aes(x = Schooling, y = LifeExp)) + 
  geom_point(size = 0.8, color = 'green') +
  geom_smooth()+
  theme_bw()+ 
  xlab('Average Schooling Year') +
  labs(title = 'Figure 5. Life Expectancy versus Formal Education, 2000-2015 (years)')

ledata %>% select(Region, Schooling, LifeExp) %>%
 group_by(Region) %>%
  summarise(Avg_Schooling = mean(Schooling),
            Avg_LifeExp = mean(LifeExp)) %>%
  ggplot(aes(x = Avg_Schooling, y = Avg_LifeExp)) +
  geom_point(size = 2.5, color = 'blue', alpha = 0.5) +
  theme_bw() +
  geom_text_repel(aes(label = Region), size = 3) +
  labs(title = 'Figure 6. Life Expectancy versus Formal Education, by Region
')

```

The graphs depict the correlation between life expectancy and the
variables adult mortality, deaths under five years old and infant
deaths. All three variables exhibit a negative relationship with life
expectancy. As the instances of adult mortality, deaths under five years
old and infant deaths per 1000 rise, life expectancy decreases
accordingly.

```{r}
ledata %>% select(Region,InfDth, AdultMort, LifeExp) %>%
  ggplot(aes(x = InfDth, y = LifeExp)) + 
  geom_point(size = 0.5, color = 'orange') +
  theme_bw() +
  labs(title = 'Figure 7. Life Expectancy versus Infant Deaths, per 1000 population')

ledata %>% select(InfDth, AdultMort, LifeExp) %>%
  ggplot(aes(x = AdultMort, y = LifeExp)) + 
  geom_point(size = 0.5, color = 'orange') +
  theme_bw() +
  labs(title = 'Figure 8. Life Expectancy versus Adult Mortality, per 1000 population')

ledata %>% 
  ggplot(aes(x = UnFiveDths, y = LifeExp)) + 
  geom_point(size = 0.5, color = 'orange') +
  theme_bw() +
  labs(title = 'Figure 9. Life Expectancy versus Children Deaths under 5 Years of Age, per 1000 population')
```

There was no data recorded for the developed regions in South Africa,
Central America and Caribbean, and Africa. There was no data recorded on
developing in the European Union region.

```{r}
ggplot(ledata, aes(x = factor(Developing), y = LifeExp, fill = factor(Developing))) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge", color = "black", width = 0.7) +
  labs(title = "Figure 10. Life Expectancy by Economy Status and Region", x = "Economy Status", y = "Mean Life Expectancy") +
  scale_fill_manual(values = c("skyblue", "salmon"), labels = c("Developing", "Developed")) +
  theme_minimal() +
  theme(legend.title = element_blank()) +
  facet_wrap(~ Region)

```

In general, a direct relationship typically exists between life
expectancy and factors such as BMI and Diphtheria immunization.
Moreover, as GDP increases, life expectancy tends to rise exponentially.
Once GDP surpasses \$30,000, the upward trend in life expectancy
continues steadily.

```{r}
ledata %>% 
  ggplot(aes(x = BMI, y = LifeExp)) + 
  geom_point(size = 0.5, color = 'orange') +
  theme_bw() +
  geom_smooth(method = lm) +
  labs(title = 'Figure 11. Life Expectancy versus Body Mass Index (BMI) of adults
')

ledata %>% 
  ggplot(aes(x = GDP, y = LifeExp)) + 
  geom_point(size = 0.5, color = 'orange') +
  theme_bw() +
  geom_smooth() +
  labs(title = 'Figure 12. Life Expectancy versus GDP (USD)')

ledata_noNA %>% 
  ggplot(aes(x = Diphtheria, y = LifeExp)) + 
  geom_point(size = 0.5, color = 'orange') +
  theme_bw() +
  geom_smooth() +
  labs(title = 'Figure 13. Life Expectancy versus DTP3 immunization coverage among 1-year-olds (%)') + xlab('DTP3 immunization rate')
```

### 3.3 Bayesian Linear Regression Analysis

#### 3.3.1 Create a BLR model

```{r}
# BLR model
glm_post1 <- stan_glm(LifeExp~., data=BLR_data, family=gaussian, seed = 1234)
glm_post1

```

In this analysis, we employ the **stan_glm()** function from the
**rstan** package to establish a Bayesian Linear Regression model. Our
dataset comprises 2,864 observations, consisting of six predictors and
one response variable.

Upon reviewing the summary, we discern parameter estimates for the
model. Notably, the model's equation can be represented as follows:

$$
LifeExp = 37.5 + 0.9×BMI + 1.2×Schooling + -1.6×HIV
$$

#### 3.3.2 Evaluate the performance of BLR model

```{r}
# Checking R hat.
summary(glm_post1)

# Assess the convergence of the MCMC chains and the behavior of the parameters during sampling
stan_trace(glm_post1, pars=c("(Intercept)", "Alcohol","GDP", "BMI",  "Schooling", "HIV","sigma")) + labs(title = 'Figure 14. Trace Plots of Each Variable')
```

$\hat{R}$ compares the variability in sampled parameter estimates across
all chains combined with the variability within each individual change.
$\hat{R} = 1$ shows stability across chains.

The **stan_trace()** function was used to visualize the traces of
parameters estimated by Markov chain Monte Carlo (MCMC) sampling. Each
trace plot to each variable appears centered around a single value and
all chains are well overlapping around a single value.

```{r}
# look at posterior predictive checks.
pp_check(glm_post1) 
```

The dark blue line shows the observed data while the light blue lines
are simulations from the posterior predictive distribution. Although the
peaks of both the observed data distribution and the posterior
distribution align along the x-axis, there's a notable discrepancy in
the density values along the y-axis.

```{r}
# View histogram of each parameter
stan_hist(glm_post1, pars=c("(Intercept)","Alcohol", "GDP", "BMI", "Schooling", "HIV","sigma"), bins=40)

# Extract posterior samples
post_samps <- as.data.frame(glm_post1, pars=c("(Intercept)","Alcohol", "GDP", "BMI",  "Schooling", "HIV","sigma"))

# calculate the median and mean of each parameter
round(apply(post_samps, 2, mean),5)
round(apply(post_samps, 2, median),5)
```

The histograms exhibit the distribution of posterior samples obtained
from the model. We also extracted posterior samples from the model and
converted to a data frame in order to calculate the the mean and median
values of each parameter.

```{r}
# Checking default priors
prior_summary(glm_post1)

# To see how the observed data has changed the parameter estimates
posterior_vs_prior(glm_post1, group_by_parameter = TRUE, pars=c("(Intercept)","Alcohol", "GDP", "BMI", "Schooling", "HIV","sigma"))
```

#### 3.3.3 Compare with Frenquentist method

```{r}
# Establish an Ordinary Linear Model
glm_fit <- glm(LifeExp~., data=BLR_data, family=gaussian)
summary(glm_fit)
round(glm_fit$coefficients, 5)
```

The maximum likelihood estimates of the intercept and slopes are nearly
identical to the mean or median values of posterior samples obtained
from Bayesian Linear Regression. Overall, the model's equation can be
expressed as follows:

$$
LifeExp = 37.5 + 0.9×BMI + 1.2×Schooling + -1.6×HIV
$$

## 4. Conclusion

In this project, we successfully explored the effects of the predictors
on life expectancy and relationships among them. There were some notable
observations while conducting our analysis. There was a significant
negative relationship between adult mortality, infants deaths, deaths
under five years old, and life expectancy. As the rates of death of
adults, infants, and children under 5 years old go up, life expectancy
tends to shorten. Additionally, the rates of thinness of adolescents and
adults are also critical factors affecting life expectancy, and there is
a negative relationship existing between them. Life expectancy varies
depending on the disparity in years of schooling across regions.
Generally, life expectancy tends to rise with increase of schooling and
higher level of economy development since 2000. Due to the presence of
correlation and multicollinearity, we selected only six variables as
predictors for conducting Bayesian Linear Regression Analysis.

In BRL analysis, our model is a perfect fit to the dataset of life expectancy
when considering the variables of GDP, BMI, alcohol, schooling and HIV.
The convergence of four chains is strong, and it's centered around a
single value, indicating the excellent performance of the model.
Moreover, through observing the BLR model, the variables of BMI,
schooling, and HIV have more weight on predicting life expectancy.

## References
