
#Load packages:
```{r}
install.packages('mice')
library('mice')
library('tidyverse')

summary(credit_data)
```


#This contains a subset of the credit data where the rows having missing value.
```{r}
rows_na = credit_data[!complete.cases(credit_data),]  

head(rows_na, 10)
```

#This shows a matrix where each row corresponds to a missing data pattern in the credit data set.
```{r}
md.pattern(credit_data)

```


#In order to perform multiple imputation on categorical data, all string variables must be converted to
# factors using as.factor()
```{r}
credit_data$Status = as.factor(credit_data$Status)
credit_data$Home = as.factor(credit_data$Home)
credit_data$Marital = as.factor(credit_data$Marital)
credit_data$Records = as.factor(credit_data$Records)
credit_data$Job = as.factor(credit_data$Job)
```



#Using the mice() function, 5 (default) multiple imputations for the null values for the credit data
# will be generated.
```{r}
Multiple_Imputation = mice(data = credit_data,  defaultMethod = c("pmm", "logreg", "polyreg", "polr"), set.seed = 1337)
```



# Show the imputed values. Columns are imputations, rows are observations
```{r}
Multiple_Imputation$imp
```

